---
title: "Practicum 1"
output: html_notebook
Author: Aadhar_Bansal, Shanmugam_Sankarakumar
email1: bansal.aad@northeastern.edu
email2: sankarakumar.s@northeastern.edu
---

```{r}
rm(list = ls())
library(RMySQL)
library(DBI)
```



```{r}
db_user <- 'admin'
db_password <- 'Dbms1234'
db_name <- 'SandboxDB'
db_host <- "cs5200fall-dbs.c6krysz4zcz9.us-east-1.rds.amazonaws.com" # AWS HOST
db_port <- 3306


mydb <- dbConnect(MySQL(), user = db_user, password = db_password,
                 dbname = db_name, host = db_host, port = db_port)
```

This Parcticum 1 is done as a pair by Aadhar_Bansal and Shanmugam_Sankarakumar. we started working on the practicum together from start till the end, we didn't splitup.

## Q1
## Drop table if exists

```{sql connection=mydb}

DROP TABLE IF EXISTS incidents

```

```{sql connection=mydb}

DROP TABLE IF EXISTS airports

```

```{sql connection=mydb}

DROP TABLE IF EXISTS conditions

```




## Creating database tables
```{sql connection=mydb}
CREATE TABLE airports (
  aid INT NOT NULL,
  airportName VARCHAR(45) NOT NULL,
  airportCode VARCHAR(45) NULL,
  state VARCHAR(45) NULL,
  PRIMARY KEY (aid));
```

#Creating lookup table conditions
```{sql connection=mydb}
CREATE TABLE `SandboxDB`.`conditions` (
  `cid` INT NOT NULL,
  `condition` VARCHAR(45) NOT NULL,
  `explanation` VARCHAR(45) NULL,
  PRIMARY KEY (`cid`));


```

creating table incidetns
```{sql connection=mydb}
CREATE TABLE incidents (
  rid INT NOT NULL,
  date VARCHAR(45) NOT NULL,
  arid INT NOT NULL,
  origin VARCHAR(45) NOT NULL,
  airline VARCHAR(45) NOT NULL,
  aircraft VARCHAR(45) NOT NULL,
  flightPhase VARCHAR(45) NOT NULL,
  altitude  INT NOT NULL,
  conditions VARCHAR(45) NULL,
  crid INT NOT NULL,
  warning VARCHAR(45) NOT NULL,
  PRIMARY KEY (rid),
  FOREIGN KEY (arid) REFERENCES airports(aid),
  FOREIGN KEY (crid) REFERENCES conditions(cid)
  );

```



## Q2 Reading the CSV File and then Processing.

```{r loadCSV}
raw_file = "BirdStrikesData-V2.csv"

df_raw <- read.csv(file = raw_file,
                   header = T,
                   stringsAsFactors = F)
```

#This code removes the military flights from the airline column
```{r Removing the Military flight}
df_raw <- subset(df_raw, df_raw$airline != "MILITARY")
num_of_rows <- nrow(df_raw)
```
#If the row values of the airport column is empty we give it unknown
```{r}
for (i in 1:num_of_rows){
  if(df_raw[i, "airport"] == "")
  {
    df_raw[i, "airport"] = "UNKNOWN"
  }
  
  if(df_raw[i, "airline"] == "")
  {
    df_raw[i, "airline"] = "UNKNOWN"
    }
}
```

#Providing data from csv to the dataframe
```{r incidents dataframe}
num_of_rows <- nrow(df_raw)
df_incidents <- data.frame(rid = df_raw$rid, 
                           date = df_raw$flight_date,
                           arid = 1,
                           origin = df_raw$origin,
                           airline = df_raw$airline,
                           aircraft = df_raw$aircraft,
                           flightPhase = df_raw$flight_phase,
                           altitude = df_raw$altitude_ft,
                           conditions = df_raw$sky_conditions,
                           warning = df_raw$pilot_warned_flag
                           
                           ) 
```

#Providing lookup tables conditions
```{r lookUp table conditions}
cond_df_raw = unique(df_incidents["conditions"])
len_cond = nrow(cond_df_raw)
df_conditions <- data.frame(cid = seq(1,len_cond),
                            condition = cond_df_raw$conditions,
                            explanation = ""
                            )
```

```{r airports}
df_airports <- data.frame(
                          airportName = df_raw$airport,
                          airportCode = "",
                          state = df_raw$origin
                            )


```

## dataframe new_df_airports, which stores the distinct of df_airports
```{r new_df_airport}
new_df_airport = unique(df_airports)
len_new_df_airport = nrow(new_df_airport)

```

## intializing the aid
```{r}
for (i in 1:len_new_df_airport){
  new_df_airport[i, "aid"] = 9 + i
}
```


## insering aid in the incidents table
```{r}
for (r in 1:num_of_rows) {
  # find the airline PK for that flight in the airlines table
    df_incidents$arid[r] <- new_df_airport$aid[new_df_airport$airportName == df_raw$airport[r]]
  
} 
```
#The warnings were neglected as we got proper output for the table incidents

## Splitting date and choosing date and neglecting time
```{r}
df_incidents$date <- substr(as.character(df_incidents$date), 1, nchar(as.character(df_incidents$date))-5)
```



## Q3 data from dataFrames

```{r}
head(df_incidents)
```

```{r}
head(new_df_airport)
```

```{r}
df_conditions
```

# Writing DataFrame back to Database
```{r Writing in Database}
dbWriteTable(mydb, "incidents", df_incidents, overwrite = T)
dbWriteTable(mydb, "airports", new_df_airport, overwrite = T)
dbWriteTable(mydb, "conditions", df_conditions, overwrite = T)


```


## Q4
#Finding top 10 airlines with greatest incidents
```{sql connection=mydb}
SELECT COUNT(row_names) as total, airline FROM incidents
GROUP BY airline
ORDER BY COUNT(row_names) desc limit 10

```

## Q5
#Displaying flight phase column which has above average no.of bird strike incidents
```{sql connection=mydb}
SELECT flightPhase, count(row_names) as Above_Average From incidents
GROUP BY flightPhase 
HAVING Above_Average > ((SELECT Count(row_names) FROM incidents)/(SELECT Count(DISTINCT flightPhase) FROM incidents))
ORDER BY Above_Average desc
```


## Q6
#Number of bird strike incidents by month
```{sql connection=mydb}
SELECT COUNT(format(date, "yyyy_mm")) as month_count, format(date, "yyyy_mm") as month FROM incidents
GROUP BY (format(date, "yyyy_mm"))

```

## Q7
#Displaying bargraph for birdstike incidents by year
```{r}
sqlCmd = "select COUNT(row_names) as count, RIGHT(date, 4) as year from incidents
 group by year
 having year >=2005 and year<=2011
 Order by year desc;";


rs = dbGetQuery(mydb, sqlCmd)

year <- c("2005","2006","2007","2008", "2009", "2010", "2011")


barplot(rs$count, names.arg = rs$year, xlab ="Year", ylab ="No of Bird Strikes", main="Bird Strikes by Year", col = "blue")

```

## Q8
```{sql connection=mydb}

DROP PROCEDURE IF EXISTS insertNEWIncident

```

## Creating Procedure
```{sql connection=mydb}
CREATE PROCEDURE insertNEWIncident(IN row_names TEXT,
        rid BIGINT,
        date TEXT,
        arid DOUBLE,
        origin TEXT,
        airline TEXT,
        aircraft TEXT,
        flightPhase TEXT,
        altitude  TEXT,
        conditions TEXT,
        warning TEXT)
BEGIN
    INSERT INTO incidents
    VALUES(row_names, rid, date, arid, origin, airline, aircraft, flightPhase, altitude, conditions,             warning);
END
```

## Calling Procedure
```{sql connection=mydb}
call insertNewIncident("1234456", 1, "10/30/2022", 9, "Delhi", "IndianAirways", "Airplane", "Climb", "0", "No Cloud","N")
```

## Checking Procedure
```{sql connection=mydb}
SELECT * from incidents
WHERE origin = "Delhi"
```


